Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs

### 2024-03-29 paper/paper.qmd
Add michaels ratings

---
title: "Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs"
author: 
    - name: Rohan Alexander^[Information and Statistical Sciences, University of Toronto. rohan.alexander@utoronto.ca.]
    - name: Lindsay Katz^[The Investigative Journalism Foundation]
    - name: Callandra Moore^[The Investigative Journalism Foundation]
    - name: Michaela Drouillard^[The Investigative Journalism Foundation]
    - name: Zane Schwartz^[The Investigative Journalism Foundation]
date: today
date-format: long
abstract: "We investigated the potential of large language models (LLMs) in developing dataset validation tests. We carried out an experiment with 96 conditions each for both GPT-3.5 and GPT-4, examining different prompt scenarios, learning modes, temperature settings, and roles. The prompt scenarios were: 1) Asking for expectations, 2) Asking for expectations with a given context, 3) Asking for expectations after requesting a data simulation, and 4) Asking for expectations with a provided data sample. For learning modes, we tested: 1) zero-shot, 2) one-shot, and 3) few-shot learning. We also tested four temperature settings: 0, 0.4, 0.6, and 1. Furthermore, two distinct roles were considered: 1) helpful assistant, 2) expert data scientist. To gauge consistency, every setup was tested five times. The LLM-generated responses were benchmarked against a gold standard suite, created by an experienced data scientist knowledgeable about the data in question. We find there are considerable returns to the use of few-shot learning, and that the more explicit the data setting can be the better. The best LLM configurations complement, rather than substitute, the gold standard results. This study underscores the value LLMs can bring to the data cleaning and preparation stages of the data science workflow."
thanks: "We thank Monica Alexander, Chris Maddison, Yadong Liu, and Will Jettinghoff for helpful suggestions. Code and data: https://github.com/RohanAlexander/evaluating_decency_and_consistency. Contributions: RA had the initial idea. RA and CM developed the experimental set-up. ZS established the political donations datasets and obtained funding. LK developed the initial suite of dataset validation tests that we compare the LLMs against. RA wrote the code to interact with the LLMs, evaluated the LLM outputs, and did the modelling. MD evaluated the LLM output. All authors contributed to writing the initial draft as well as improving and finalizing the paper."
bibliography: "bibliography.bib"
format:
  pdf:
    keep-tex: false
header-includes:
  - \usepackage{listings}
  - \lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
number-sections: true
---

@@

The Investigative Journalism Foundation (IJF) created and maintains a dataset related to political donations in Canada. As of September 2023, the dataset comprises 9,204,112 observations and 15 variables. Every day new observations are added, based on newly released donations records made available by provincial and federal elections agencies. This release cycle is variable, with periods of inactivity followed by bursts of multiple releases. @katzandmoore manually construct an extensive suite of automated tests for this dataset. These impose certain minimum standards on the dataset, including: that constituent aspects add to match any total, class is appropriate, and null values are where expected. This suite allows researchers to use the dataset with confidence and ensures that new additions are fit for purpose.

We revisit this suite of tests to determine whether we can use large language models (LLMs) to mimic this suite of validation tests. We consider a variety of prompts, roles, learning modes, and temperature settings, resulting in an experiment with 96 conditions. In particular, we consider four prompt variations: ask for expectations; ask for expectations given described context; ask for expectations having first asked for a data simulation; and ask for expectations given a sample of data. We also consider zero-, one-, and few-shot learning; four temperature values: 0, 0.4, 0.6, and 1 (which should influence how unexpected the results are); and two roles: helpful assistant and expert data scientist. For every combination we obtain five responses from the LLM. We run all conditions separately for both GPT-3.5 and GPT-4.

@@

Historically, natural language processing (NLP) models operated in much smaller contexts than today. The first NLP systems were rule-based, relying on hand-crafted linguistic rules based on experts\' understanding of languages, and were trained on much smaller datasets.

@@

The transition from statistical methods and traditional machine learning methods to \"deep learning\" came when @bengio2003 introduced neural-network based language models. These models predicted the next word in a sentence, given the previous words. Bengio et al\'s model relied on a representation for words, where every word from the vocabulary is linked with a continuous vector. During training, models refined these vector representations. Words with similar meanings had vectors that were closer together in vector space, which enables the model to understand semantic relationships between words.

Once neural networks were introduced, many significant advancements such as word embeddings ([@pennington2014], [@mikolov2013]), long short-term memory  [@hochreiter1997] and Convolutional Neural Networks (CNNs) [@kim2014] followed.  The concept of pre-training models and then fine-tuning them for specific tasks (as seen with models like ULMFiT [@howard2018] and ELMo [@peters2018]) started to take root in the late 2010s, which set the stage for transformer-based architectures.

@@

We are interested in the extent to which the LLMs can develop a suite of data validation tests that is similar to a suite developed by an experienced expert data scientist who is familiar with the dataset. To test this, we establish and run a series of experiments where we consider different specifications and then compare the LLM output. In particular, the variables that we consider are briefly summarized here, and the full details are included in Appendix @sec-details.

- Four prompts:
    - List the variables of interest -  "index", "amount", "donor_location", "donation_date", "donor_full_name", "donor_type", "political_entity", "political_party", "recipient", "region", "donation_year", "amount_monetary", "amount_non_monetary", "electoral_event", "electoral_district", "added" - and ask for a suite of expectations using the Python package great_expectations.
    - As above, plus provide details about each variable. For instance, "amount" is a monetary value that cannot be less than $0. An example observation is "195.46". It is, possible, but unlikely to be more than $1,000.00. It cannot be NA. It should be a numeric. The maximum donation "amount" depends on the value of "region" and "year". For "Federal" is 1675, for "Quebec" is 100 since 2013 and 500 for earlier years, for "British Columbia" is 1309.09, for "Ontario" is 3325, and for "Alberta" is 4300. There is no limit for "Saskatchewan".
    - As above, plus ask the LLM to simulate an example dataset of 1000 observations.
    - Provide the details of the dataset in the first option, as well as an extract of the dataset.
- Three learning modes: zero-, one-, and few-shot learning. In the first, we do not provide any examples; in the second, we provide one example; and in the third, we provide three examples.
- Four temperature values: 0, 0.4, 0.6, and 1.
- Two roles:
    - You are a helpful assistant.
    - You are a highly-trained, experienced, data scientist who is an expert at writing readable, correct, Python code.

@@

This results in a dataset of responses. The combination of settings that gave rise to each response was blinded and the order randomized, and then the responses were ranked by two experienced human evaluators on two metrics.

@@

library(tidyverse)

data <- read_csv("../raw_data/responses_with_human_coding.csv")

data <- 
    data |>
    mutate(Prompt_n = factor(Prompt_n, levels = c("Name", "Describe", "Simulate", "Example")),
           Temperature = factor(Temperature),
           Role_n = factor(Role_n, levels = c("Helpful", "Expert")),
           Shot_n = factor(Shot_n, levels = c("Zero", "One", "Few")),
           Version = factor(Version)
    )

data <- 
    data |>
    rowwise() |>
    mutate(
        consistency = round(mean(c(consistency_coder_1, consistency_coder_2), na.rm = TRUE)),
        decency = round(mean(c(decency_coder_1, decency_coder_2), na.rm = TRUE))) |>
    ungroup()
```

@fig-intercoder examines how the two evaluators rated decency and consistency. In particular, @fig-intercoder-1 focuses on how they agreed on decency, and @fig-intercoder-2 shows how they agreed on consistency. If both evaluators rated every response the same, then we would expect all responses to be on the 45 degree line. We do not find many absolute differences of opinion, but there nonetheless is still a wide range of differences, especially in terms of decency.

data |> 
    ggplot(aes(x = consistency_coder_1, y = consistency_coder_2)) +
    geom_jitter(width = 0.15, height = 0.15, alpha = 0.5) +
    theme_minimal() +
    labs(
        x = "Evaluator 1", y = "Evaluator 2"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

To bring the two evaluators together we average their response and then round it to the nearest integer. This then allows us to examine whether there is a relationship between decency and consistency (@fig-decencyvsconsistency).

@@

data |> 
    ggplot(aes(x = decency, y = consistency)) +
    geom_jitter(width = 0.15, height = 0.15, alpha = 0.5) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Consistency"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

@fig-version examines how decency and consistency differ based on whether GPT-3.5 or GPT-4 is used. Unexpectedly, GPT-4 has fewer responses rated 5/5 for decency, compared with GPT-3.5 (@fig-version-1). GPT-3.5 has fewer responses rated 2/5, but overall the mean decency response for GPT-3.5 is 3.23, while the mean decency of the responses generated by GPT-4 is 3.01. The consistency is not too different between the two versions, with GPT-3.5 having an average of 3.61, while GPT-4 has an average of 3.65. GPT-4 appears to have fewer responses that are completely identical (@fig-version-2).

@@

data |> 
    ggplot(aes(x = decency, fill = Version)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency, fill = Version)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

@@

There appears to be considerable difference in terms of how the prompts are associated with decency. In particular, "Name" is never associated with a 5/5 rating (@fig-prompt-1). Surprisingly, however, the most informative prompt, "Example", is also never associated with a 5/5 rating. Instead it is "Describe" and "Simulate" that tend to be associated with better decency ratings. This is reflected in the averages, which are 2.65, 3.46, 3.44, and 2.94 for "Name", "Describe", "Simulate", and "Example", respectively.

The pattern is not as clear when it comes to consistency (@fig-prompt-2). All four have similar averages, at 3.71, 3.75, 3.48, and 3.58 for "Name", "Describe", "Simulate", and "Example", respectively. That said, it is clear that a wider variety of responses (as denoted by lower consistency ratings), are rarely seen for "Name" and "Describe".

@@

data |>
    group_by(Prompt_n) |>
    summarize(mean_decency = mean(decency),
             mean_consistency = mean(consistency)
                )
```

@@

data |> 
    ggplot(aes(x = decency, fill = Prompt_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences", fill = "Prompt type"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency, fill = Prompt_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences", fill = "Prompt type"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

@@

@fig-temperature examines how decency and consistency differ based on which of the four temperature values we consider---0, 0.4, 0.6, 1---is used. There appears to be limited difference in terms of how different temperature values are associated with decency (@fig-temperature-1). They all have similar mean values at 3.10, 3.25, 2.98, and 3.15 for temperature values of 0, 0.4, 0.6, and 1, respectively.

In contrast, as expected temperature has an effect on consistency. Temperature values of 0 are associated with ratings of high consistency, and higher temperatures are associated with lower consistency (@fig-temperature-2). Their means differ considerably, with 4.77, 3.75, 3.31, and 2.69 for temperature values of 0, 0.4, 0.6, and 1, respectively.

@@

data |>
    group_by(Temperature) |>
    summarize(mean_decency = mean(decency),
             mean_consistency = mean(consistency)
                )
```

@@

data |> 
    ggplot(aes(x = decency, fill = Temperature)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency, fill = Temperature)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

Role is an aspect of a prompt that is provided to the LLM before the main prompt content. We considered two different roles, one that positioned the LLM as a helpful assistant, and the other that positioned the LLM as an experienced data scientist (@fig-role). We were expecting that the expert role would result in better code, but there was no obvious difference in terms of decency (@fig-role-1) or consistency (@fig-role-2). Their means did not differ by much in either case.

@@

data |>
    group_by(Role_n) |>
    summarize(mean_decency = mean(decency),
             mean_consistency = mean(consistency)
                )
```

```{r}
#| echo: false
#| label: fig-role
#| fig-cap: "How decency and consistency change based on whether role is 'Helpful' or 'Expert'"
#| fig-subcap: ["Decency","Consistency"]
#| warning: false
#| layout-ncol: 2
#| 
data |> 
    ggplot(aes(x = decency, fill = Role_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences", fill = "Role"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency, fill = Role_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences", fill = "Role"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

@@

We find substantial differences, especially when moving away from zero-shot learning (@fig-shot). In particular, we see that decency of 1/5 is dominated by zero-shot, while zero-shot is under-represented in 5/5 (@fig-shot-1). This is also reflected in the mean decency which for zero-shot is 2.83, while for one- and few-shot learning is 3.34 and 3.19, respectively.

We see this pattern in consistency as well. For instance, zero-shot learning is over-represented in the least consistent responses, both 1/5 and 2/5 (@fig-shot-2). And the mean level of consistency is lower for zero-shot learning, at 3.45, compared with single- and few-shot, at 3.77 and 3.67, respectively.

@@

data |>
    group_by(Shot_n) |>
    summarize(mean_decency = mean(decency),
             mean_consistency = mean(consistency)
                )
```

@@

data |> 
    ggplot(aes(x = decency, fill = Shot_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences", fill = "Shots"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency, fill = Shot_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences", fill = "Shots"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

>>>>

---
title: "Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs"
subtitle: "An application to Canadian political donations data"
author: 
    - name: Rohan Alexander^[Information and Statistical Sciences, University of Toronto. rohan.alexander@utoronto.ca.]
    - name: Lindsay Katz^[The Investigative Journalism Foundation]
    - name: Callandra Moore^[The Investigative Journalism Foundation]
    - name: Michaela Drouillard^[The Investigative Journalism Foundation]
    - name: Michael Wing-Cheung Wong^[Massachusetts Institute of Technology]
    - name: Zane Schwartz^[The Investigative Journalism Foundation]
date: today
date-format: long
abstract: "We investigated the potential of large language models (LLMs) in developing dataset validation tests. We carried out an experiment with 96 conditions each for both GPT-3.5 and GPT-4, examining different prompt scenarios, learning modes, temperature settings, and roles. The prompt scenarios were: 1) Asking for expectations, 2) Asking for expectations with a given context, 3) Asking for expectations after requesting a data simulation, and 4) Asking for expectations with a provided data sample. For learning modes, we tested: 1) zero-shot, 2) one-shot, and 3) few-shot learning. We also tested four temperature settings: 0, 0.4, 0.6, and 1. Furthermore, two distinct roles were considered: 1) helpful assistant, 2) expert data scientist. To gauge consistency, every setup was tested five times. The LLM-generated responses were benchmarked against a gold standard suite, created by an experienced data scientist knowledgeable about the data in question. We find there are considerable returns to the use of few-shot learning, and that the more explicit the data setting can be the better. The best LLM configurations complement, rather than substitute, the gold standard results. This study underscores the value LLMs can bring to the data cleaning and preparation stages of the data science workflow."
thanks: "We thank Monica Alexander, Chris Maddison, Dianne Cook, Yadong Liu, and Will Jettinghoff for helpful suggestions. Code and data: https://github.com/RohanAlexander/evaluating_decency_and_consistency. Contact: rohan.alexander@utoronto.ca. Contributions: RA had the initial idea. RA and CM developed the experimental set-up. ZS established the political donations datasets and obtained funding. LK developed the initial suite of dataset validation tests that we compare the LLMs against. RA wrote the code to interact with the LLMs and did the modelling. RA, MD, and MW evaluated the LLM output. All authors contributed to writing the initial draft as well as improving and finalizing the paper."
bibliography: "bibliography.bib"
format:
  pdf:
    keep-tex: false
header-includes:
  - \usepackage{listings}
  - \lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
number-sections: true
---

library(tidyverse)
library(arrow)
library(rstanarm)

data <- read_parquet("../data/cleaned/analysis_data.parquet")
```

@@

The Investigative Journalism Foundation (IJF) created and maintains a dataset related to political donations in Canada. As of September 2023, the dataset comprised 9,204,112 observations and 15 variables. Every day new observations are added, based on newly released donations records made available by provincial and federal elections agencies. This release cycle is variable, with periods of inactivity followed by bursts of multiple releases. @katzandmoore manually construct an extensive suite of automated tests for this dataset. These impose certain minimum standards on the dataset, including: that constituent aspects add to match any total, class is appropriate, and null values are where expected. This suite allows researchers to use the dataset with confidence and ensures that new additions are fit for purpose.

We revisit this suite of tests to determine whether we can use large language models (LLMs) to mimic this suite of validation tests. Our estimand is the similarity between those tests written by an LLM and those written by an experienced data scientist. We consider a variety of prompts, roles, learning modes, and temperature settings, resulting in an experiment with 96 conditions. In particular, we consider four prompt variations: ask for expectations; ask for expectations given described context; ask for expectations having first asked for a data simulation; and ask for expectations given a sample of data. We also consider zero-, one-, and few-shot learning; four temperature values: 0, 0.4, 0.6, and 1 (which should influence how unexpected the results are); and two roles: helpful assistant and expert data scientist. For every combination we obtain five responses from the LLM. We run all conditions separately for both GPT-3.5 and GPT-4.

@@

Historically, natural language processing (NLP) models operated in much smaller contexts than today. The first NLP systems were rule-based, relying on hand-crafted linguistic rules based on experts' understanding of languages, and were trained on much smaller datasets.

@@

The transition from statistical methods and traditional machine learning methods to \"deep learning\" came when @bengio2003 introduced neural-network based language models. These models predicted the next word in a sentence, given the previous words. The model of @bengio2003 relied on a representation for words, where every word from the vocabulary is linked with a continuous vector. During training, models refined these vector representations. Words with similar meanings had vectors that were closer together in vector space, which enables the model to understand semantic relationships between words.

Once neural networks were introduced, many significant advancements such as word embeddings [@pennington2014; @mikolov2013], long short-term memory [@hochreiter1997] and Convolutional Neural Networks (CNNs) [@kim2014] followed. The concept of pre-training models and then fine-tuning them for specific tasks (as seen with models like ULMFiT [@howard2018] and ELMo [@peters2018]) started to take root in the late 2010s, which set the stage for transformer-based architectures.

@@

We are interested in the extent to which the LLMs can develop a suite of data validation tests that is similar to a suite developed by an experienced expert data scientist who is familiar with the dataset. To test this, we establish and run a series of experiments where we consider different specifications and then compare the LLM output. In particular, the variables that we consider are summarized here, and the full details are included in [Appendix -@sec-details].

- Four prompts:
    - List the variables of interest -  "index", "amount", "donor_location", "donation_date", "donor_full_name", "donor_type", "political_entity", "political_party", "recipient", "region", "donation_year", "amount_monetary", "amount_non_monetary", "electoral_event", "electoral_district", "added" - and ask for a suite of expectations using the Python package `great_expectations`.
    - As above, plus provide details about each variable. For instance, "amount" is a monetary value that cannot be less than $0. An example observation is "195.46". It is, possible, but unlikely to be more than $1,000.00. It cannot be NA. It should be a numeric. The maximum donation "amount" depends on the value of "region" and "year". For "Federal" is 1675, for "Quebec" is 100 since 2013 and 500 for earlier years, for "British Columbia" is 1309.09, for "Ontario" is 3325, and for "Alberta" is 4300. There is no limit for "Saskatchewan".
    - As above, plus ask the LLM to simulate an example dataset of 1000 observations.
    - Provide the details of the dataset in the first option, as well as an extract of the dataset.
- Three learning modes: zero-, one-, and few-shot learning. In the first, we do not provide any examples; in the second, we provide one example; and in the third, we provide three examples.
- Four temperature values: 0, 0.4, 0.6, and 1.
- Two roles:
    - You are a helpful assistant.
    - You are a highly-trained, experienced, data scientist who is an expert at writing readable, correct, Python code.

@@

This results in a dataset of responses. The combination of settings that gave rise to each response was blinded and the order randomized, and then the responses were ranked by three experienced human evaluators on two metrics.

@@

To bring the three evaluators together we consider the median response. This then allows us to examine whether there is a relationship between decency and consistency (@fig-decencyvsconsistency). (Details about coder-specific responses are provided in [Appendix -@sec-intercoder].)

@@

data |>
  ggplot(aes(x = decency_median, y = consistency_median)) +
  geom_jitter(width = 0.15,
              height = 0.15,
              alpha = 0.5) +
  theme_minimal() +
  labs(x = "Decency", y = "Consistency") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@fig-version examines how decency and consistency differ based on whether GPT-3.5 or GPT-4 is used. Unexpectedly, GPT-4 has fewer responses rated 5/5 for decency, compared with GPT-3.5 (@fig-version-1). GPT-3.5 has fewer responses rated 2/5, but overall the mean decency response for GPT-3.5 is 3.26, while the mean decency of the responses generated by GPT-4 is 3.19. The consistency is not too different between the two versions, with GPT-3.5 having an average of 3.21, while GPT-4 has an average of 3.33. GPT-4 appears to have fewer responses that are completely identical (@fig-version-2).

data |>
    group_by(Version) |>
    summarize(mean_decency = mean(decency_median) |> round(2),
             mean_consistency = mean(consistency_median) |> round(2)
                )
```

@@

data |> 
  count(Version, decency_median) |> 
  group_by(Version) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(decency_median = as_factor(decency_median)) |> 
  ggplot(aes(x = Version, y = percentage, fill = decency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Decency", x = "Version", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)

data |> 
  count(Version, consistency_median) |> 
  group_by(Version) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(consistency_median = as_factor(consistency_median)) |> 
  ggplot(aes(x = Version, y = percentage, fill = consistency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Consistency", x = "Version", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)

data |> 
    ggplot(aes(x = decency_median, fill = Version)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency_median, fill = Version)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

@@

There appears to be considerable difference in terms of how the prompts are associated with decency. In particular, "Name" is never associated with a 5/5 rating (@fig-prompt-1). Surprisingly, however, the most informative prompt, "Example", is also never associated with a 5/5 rating. Instead it is "Describe" and "Simulate" that tend to be associated with better decency ratings. This is reflected in the averages, which are 3.15, 3.31, 3.27, and 3.17 for "Name", "Describe", "Simulate", and "Example", respectively.

The pattern is not as clear when it comes to consistency (@fig-prompt-2). All four have similar averages, at 3.42, 3.10, 3.21, and 3.35 for "Name", "Describe", "Simulate", and "Example", respectively. That said, it is clear that a wider variety of responses (as denoted by lower consistency ratings), are rarely seen for "Name" and "Describe".

@@

data |>
    group_by(Prompt_n) |>
    summarize(mean_decency = mean(decency_median) |> round(2),
             mean_consistency = mean(consistency_median) |> round(2)
                )
```

@@

data |> 
  count(Prompt_n, decency_median) |> 
  group_by(Prompt_n) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(decency_median = as_factor(decency_median)) |> 
  ggplot(aes(x = Prompt_n, y = percentage, fill = decency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Decency", x = "Prompt", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)

data |> 
  count(Prompt_n, consistency_median) |> 
  group_by(Prompt_n) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(consistency_median = as_factor(consistency_median)) |> 
  ggplot(aes(x = Prompt_n, y = percentage, fill = consistency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Consistency", x = "Prompt", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)

data |> 
    ggplot(aes(x = decency_median, fill = Prompt_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences", fill = "Prompt type"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency_median, fill = Prompt_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences", fill = "Prompt type"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

@@

@fig-temperature examines how decency and consistency differ based on which of the four temperature values we consider---0, 0.4, 0.6, 1---is used. There appears to be limited difference in terms of how different temperature values are associated with decency (@fig-temperature-1). They all have similar mean values at 3.25, 3.35, 2.96, and 3.33 for temperature values of 0, 0.4, 0.6, and 1, respectively.

In contrast, as expected temperature has an effect on consistency. Temperature values of 0 are associated with ratings of high consistency, and higher temperatures are associated with lower consistency (@fig-temperature-2). Their means differ considerably, with 4.73, 3.40, 2.69, and 2.27 for temperature values of 0, 0.4, 0.6, and 1, respectively.

@@

data |>
    group_by(Temperature) |>
    summarize(mean_decency = mean(decency_median) |> round(2),
             mean_consistency = mean(consistency_median) |> round(2)
                )
```

@@

data |> 
  count(Temperature, decency_median) |> 
  group_by(Temperature) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(decency_median = as_factor(decency_median)) |> 
  ggplot(aes(x = Temperature, y = percentage, fill = decency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Decency", x = "Temperature", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)

data |> 
  count(Temperature, consistency_median) |> 
  group_by(Temperature) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(consistency_median = as_factor(consistency_median)) |> 
  ggplot(aes(x = Temperature, y = percentage, fill = consistency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Consistency", x = "Temperature", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)

data |> 
    ggplot(aes(x = decency_median, fill = Temperature)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency_median, fill = Temperature)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

Role is an aspect of a prompt that is provided to the LLM before the main prompt content. We considered two different roles, one that positioned the LLM as a helpful assistant, and the other that positioned the LLM as an experienced data scientist (@fig-role). We were expecting that the expert role would result in better code, but there was no obvious difference in terms of decency (@fig-role-1) or consistency (@fig-role-2). Their mean decency did not differ by much in either case with 3.12 and 3.32 for 'Helpful' and 'Expert', respectively, while mean consistency was 3.35 and 3.19.

@@

data |>
    group_by(Role_n) |>
    summarize(mean_decency = mean(decency_median) |> round(2),
             mean_consistency = mean(consistency_median) |> round(2)
                )
```

```{r}
#| echo: false
#| label: fig-role
#| fig-cap: "How decency and consistency change based on whether role is 'Helpful' or 'Expert'"
#| fig-subcap: ["Decency","Consistency"]
#| warning: false
#| layout-ncol: 2

data |> 
  count(Role_n, decency_median) |> 
  group_by(Role_n) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(decency_median = as_factor(decency_median)) |> 
  ggplot(aes(x = Role_n, y = percentage, fill = decency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Decency", x = "Role", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)

data |> 
  count(Role_n, consistency_median) |> 
  group_by(Role_n) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(consistency_median = as_factor(consistency_median)) |> 
  ggplot(aes(x = Role_n, y = percentage, fill = consistency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Consistency", x = "Role", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)

data |> 
    ggplot(aes(x = decency_median, fill = Role_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Decency", y = "Number of occurrences", fill = "Role"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")

data |> 
    ggplot(aes(x = consistency_median, fill = Role_n)) +
    geom_histogram(position = "dodge2", binwidth = 1) +
    theme_minimal() +
    labs(
        x = "Consistency", y = "Number of occurrences", fill = "Role"
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme(legend.position = "bottom")
```

@@

We find substantial differences, especially when moving away from zero-shot learning (@fig-shot). In particular, we see that decency of 1/5 is dominated by zero-shot, while zero-shot is under-represented in 5/5 (@fig-shot-1). This is also reflected in the mean decency which for zero-shot is 2.83, while for one- and few-shot learning is 3.39 and 3.45, respectively.

We see this pattern in consistency as well. For instance, zero-shot learning is over-represented in the least consistent responses, both 1/5 and 2/5 (@fig-shot-2). And the mean level of consistency is lower for zero-shot learning, at 2.98, compared with single- and few-shot, at 3.42 and 3.41, respectively.

@@

data |>
    group_by(Shot_n) |>
    summarize(mean_decency = mean(decency_median) |> round(2),
             mean_consistency = mean(consistency_median) |> round(2)
                )
```

@@

data |> 
  count(Shot_n, decency_median) |> 
  group_by(Shot_n) |> 
  mutate(percentage = n / sum(n) * 100) |> 
  ungroup() |> 
  mutate(decency_median = as_factor(decency_median)) |> 
  ggplot(aes(x = Shot_n, y = percentage, fill = decency_median)) +
  geom_bar(position = "stack", stat = "identity", width = 0.6) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(fill = "Decency", x = "Shot", y = "Percentage (%)") +
  theme(legend.position = "bottom") +
  geom_text(aes(label = sprintf("%d%%", round(percentage))), position = position_stack(vjust = 0.5), size = 3)
